---
title: "A Data Scientist Perspective On \n Philadelphia Crime 2006-2016"
author: "CF Hainey"
date: "November 29, 2016"
geometry: margin = 2cm
output: 
  pdf_document
---

```{r global_options, include =FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE, cache = TRUE)
```

**Background**  
The open data Philly initiative has crime data from Jan 1, 2006 to Nov 11, 2016. There are over 2 million crimes events in the file.  The types of crime entries vary from disorderly conduct to criminal homicide.  Each entry has location, street address with longitude and latitude coordinates, dispatch date and time, district number, Dc-key, UCR-General and Text-General-Code. Text-General-Code is a factor list of 33 types of crimes and has corresponding Dc-key.  These two fields describe the type of crime commited.

Each crime type was calculated by the number of reported instances. Each type of crime was grouped into 4 segments: Part I, Violent and Property Crimes; Part II; and Quality of Life. Part I crimes (felony and misdemeanor) make up the Index Crimes tracked by the FBI and include Violent Crimes such as Homicide, Rape, Aggravated Assault and Robbery. Part I crimes also include the Property Crimes of Theft, Burglary, and Arson. All other misdemeanor crimes are filed under Part II crimes and include Other Assaults, Receiving Stolen Property, Embezzlement, Fraud, Weapons Violations, Forgery/Counterfeiting, Drug Law Violations, Offenses against Family, DUI, and Vandalism. Quality of Life Crimes are less severe summary offenses that affect standards of living in communities, such as Drunkenness, Prostitution, Gambling Violations, Disorderly Conduct, and other minor offenses. All Other Offenses have been grouped with Quality of Life and may include Traffic and Parking Tickets, Truancy and Curfew violations, and police investigations such as pedestrian stops and vehicle investigations.    

Table 1 describes the crimes in philadelphia with crime segments and the frequency between Jan 1, 2006 and Nov 12, 2016.  It is unfortunet that ~20% of the data is characterized as "all other offenses" and we are not able to precisely say what types of crimes are happening.  These are likely quality of life crimes and we can at least separate them from the more violent crimes.  

```{r, warning=FALSE,echo=FALSE,message=FALSE, cache=TRUE}
## data from open Philly and Kaggle.

library(data.table)
#suppressMessages(library(data.table, warn.conflicts = FALSE, quietly=TRUE))
dat <- read.csv("C:/Users/Haineycf/Documents/Crime_Philadelphia/crime.csv")
#dat <- fread("C:/Users/Haineycf/Documents/Crime_Philadelphia/crime.csv")
### fread has produced messages in pdf, even with message = FALSE
```

```{r}
### make column for month and another for year
dat$Dispatch_Date <- as.Date(dat$Dispatch_Date)
dat$year <- format(as.Date(dat$Dispatch_Date,
                           format="%m/%d/%Y"),"%Y")
dat$Month <- format(as.Date(dat$Dispatch_Date,
                            format="%m/%d/%Y"),"%m")
dat$day <- weekdays(as.Date(dat$Dispatch_Date))
#head(dat)
dat$day <- factor(dat$day, levels=c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))

## group crimes into four types: violent, partII, property and quality of life
dat$crime_type[dat$Text_General_Code == "Homicide - Gross Negligence" |
                 dat$Text_General_Code =="Rape" |
                 dat$Text_General_Code =="Robbery No Firearm" |
                 dat$Text_General_Code =="Aggravated Assault Firearm" |
                 dat$Text_General_Code =="Homicide - Justifiable" |
                 dat$Text_General_Code =="Homicide - Criminal" |
                 dat$Text_General_Code =="Aggravated Assault No Firearm" |
                 dat$Text_General_Code =="Robbery Firearm" ] <-"Violent_crime"


dat$crime_type[dat$Text_General_Code == "Burglary Residential" |
                 dat$Text_General_Code =="Burglary Non-Residential" |
                 dat$Text_General_Code =="Thefts" |
                 dat$Text_General_Code =="Theft from Vehicle" |
                 dat$Text_General_Code =="Motor Vehicle Theft" |
                 dat$Text_General_Code =="Arson" ] <-"Property"


dat$crime_type[dat$Text_General_Code == "Receiving Stolen Property"  |
                 dat$Text_General_Code =="Embezzlement" |
                 dat$Text_General_Code =="Fraud" |
                 dat$Text_General_Code =="Weapon Violations" |
                 dat$Text_General_Code =="Forgery and Counterfeiting" |
                 dat$Text_General_Code =="Narcotic / Drug Law Violations" |
                 dat$Text_General_Code =="Offenses Against Family and Children" |
                 dat$Text_General_Code =="DRIVING UNDER THE INFLUENCE"  |
                 dat$Text_General_Code =="Other Assaults"  |  ### does this belong here?
                 dat$Text_General_Code =="Vandalism/Criminal Mischief" ] <-"PartII"

dat$crime_type[dat$Text_General_Code == "Disorderly Conduct"  |
                 dat$Text_General_Code =="All Other Offenses" |
                 dat$Text_General_Code =="Vagrancy/Loitering" |
                 dat$Text_General_Code =="Other Sex Offenses (Not Commercialized)" |
                 dat$Text_General_Code =="Liquor Law Violations" |
                 dat$Text_General_Code =="Gambling Violations" |
                 dat$Text_General_Code =="Prostitution and Commercialized Vice"  |
                 dat$Text_General_Code ==""  |
                 dat$Text_General_Code =="Public Drunkenness" ] <-"Quality_of_Life"


### create table of crime and the type of crime it is
crime_occurance <-table(dat$Text_General_Code, dat$crime_type)
#head(crime_occurance)
df <-as.data.frame(crime_occurance)
df <-df[order(df$Freq, decreasing = TRUE),]

df <-df[ which(df$Freq >0), ]
row.names(df) <- NULL

# colnames is not working with knit pdf!?? not sure why.
#colnames(df)<-c("Text_General_Code","Crime Segment","Freq")


## Pander formats data nicely into table
library(pander)
pander(df)


### sum different variables
t_violent <-with(df, sum(df[Var2 =="Violent_crime", "Freq"]))
t_assault <-with(df, sum(df[Var1 =="Aggravated Assault No Firearm", "Freq"]))
a_crime <- t_assault/2182727

t_PartII <-with(df, sum(df[Var2 =="PartII", "Freq"]))

t_Quality_of_Life <-with(df, sum(df[Var2 =="Quality_of_Life", "Freq"]))

ql_all_others <-with(df, sum(df[Var1 =="All Other Offenses", "Freq"]))
t_Property <-with(df, sum(df[Var2 =="Property", "Freq"]))
theft_property <-with(df, sum(df[Var1 =="Thefts", "Freq"]))
vand <-with(df, sum(df[Var1 =="Vandalism/Criminal Mischief", "Freq"]))

```

```{r, include=FALSE, eval=FALSE}
### this chunk evaluates which day most crime occured

most.crimecrime_occurance <-table(dat$Dispatch_Date)
most.crimecrime_occurance <-as.data.frame(most.crimecrime_occurance)
most.crimecrime_occurance <-most.crimecrime_occurance[order(most.crimecrime_occurance$Freq, decreasing = TRUE),]
tail(most.crimecrime_occurance)

```
**Results**  
For the decade spanning 2006-2016, Part II crimes make up `r (t_PartII/2182727)*100`% of all crime commited from Jan 2006 to Nov 2016, Property is `r (t_Property/2182727)*100`% of all crime, `r (t_Quality_of_Life/2182727)*100`% represent quality of life crimes and the last `r (t_violent/2182727)*100`% represent violent crimes.  The crime most committed was "All other Offenses" and is `r (ql_all_others/2182727)*100`% of total crime. Homicide- Gross Negligence had 12 insteances or 0.0005% of all crime.   The following pie charts goes into more detail of the 4 segments of crime and what types of crime.  

```{r, fig.width=7, fig.height=5, fig.path='Figs/'}

#### look at each type of crime and make a pie of it
crime.type <-table(dat$Text_General_Code, dat$crime_type)
crime.type <-as.data.frame(crime.type)
violent <- subset(crime.type,Var2 == "Violent_crime" & Freq >0)
quality <- subset(crime.type,Var2 == "Quality_of_Life"& Freq >0)
property <- subset(crime.type,Var2 == "Property"& Freq >0)
partII <- subset(crime.type,Var2 == "PartII"& Freq >0)

par(mfrow = c(2,2),mar=c(1, 1, 2, 11),oma=c(0,0,0,0), xpd=TRUE)
## mar is bottom, left, top, right
pie(x=violent$Freq, label="", col=rainbow(length(violent$Var1)), main="Violent\n Crimes" ) 
legend("topright", inset=c(-1,.2),legend=violent$Var1, fill=rainbow(length(violent$Var1)), cex = 0.5)

#par(mar=c(5.1, 4.1, 4.1, 10), xpd=TRUE)
pie(x=quality$Freq, label="", col=rainbow(length(quality$Var1)), main="Quality of life \n Crimes" ) 
legend("topright", inset=c(-1.2,.2),legend=quality$Var1, fill=rainbow(length(quality$Var1)), cex = 0.5)

#par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
pie(x=property$Freq, label="", col=rainbow(length(property$Var1)), main="property \n Crimes" ) 
legend("topright", inset=c(-1,.2),legend=property$Var1, fill=rainbow(length(property$Var1)), cex = 0.5)

#par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
pie(x=partII$Freq, label="", col=rainbow(length(partII$Var1)), main="partII\n Crimes" ) 
legend("topright", inset=c(-1.2,.2),legend=partII$Var1, fill=rainbow(length(partII$Var1)), cex = 0.5)

```

    
```{r, fig.width=6, fig.height=4, fig.path='Figs/'}
## create table related to time: year, month, day, hour
crime_yr <-as.data.frame(table(dat$year))
crime_mn <-as.data.frame(table(dat$Month))
crime_day <-as.data.frame(table(dat$day))
crime_hr <-as.data.frame(table(dat$Hour))
```

**Crime With Respect To Time**  
Aggregating time with crime segments reveals broad trends of crimes.  In 2006, Philadelphia had 234,755 criminal events and 183,300 in 2015.  This ten years span saw a decline in crime rate of 22% in Philadelphia.  Even though, all crime segments have declined from 2006 to 2015, not all crime segments have declined year after year.  Violent crimes do show a year over year decrease, with 22,198 in 2006 to 15,768 in 2015, a `r ((22198-15768)/22198)*100`% reduction.  Part 2 crimes show the same trend ranging from 54,991 and 41,652 in 2015 or `r ((54991-41652)/54991)*100`% reduction.  

Property crimes also reduced from 59,772 in 2006 to 40,260 in 2016 `r ((59772-40260)/59772)*100`% reduction.  However, there was an increase in property crime from 2009 to 2011, 51,580 and 59,335 incidence respectfully, this produced a `r ((59335-51580)/51580)*100`% increase in a short amount of time. Quality of life also has a blip.  In 2006 there were 59,038 reports and dips down to 39,744 in 2011 (a decrease of `r ((59038-39744)/59038)*100`%),  but its back up to 48,413 in 2015, an increase of `r ((48413-39744)/39744)*100`% from the 2011 low, but a decrease of `r ((59038-48413)/59038)*100`% from its 2006 high.  This increase could be due to many reasons.   Quality of life crimes might be increasing due to increase enforcement prioritization by District Captains.  

Other noteworthy treands show the summer months have more crimes committed.  The months of February, November and December have least amount of crime.  During the weekdays, part II has the highest reporting with a peak on Tuesday and decreases towards the weekend and into Monday.  Monday is the peak for Property crimes, but like part II, it decreases moving on into the week and weekend.  Quality of life on the other hand, has the least amount of incidence on Monday, and has the most on Saturdays.   

Analysis over the 24 hour shows that quality of life is reported most in the evening.  Some of these will be curfews, drunkenness etc.  Property crimes happen most when no one is home during school and work hours.  The least amount of crime occurs between 4am and 6am.  
 
 
```{r}

############## check to see if a particular crime is increaseing or decreasing by potting
library(ggplot2)
library(grid)
library(gridExtra)


crime_by_year <- aggregate(Month ~crime_type + year, data = dat, FUN = length)



############## crime by month
crime_by_month <- aggregate(year ~crime_type + Month,
                           data = dat, FUN = length)
library(tidyr)

data_wide <- spread(crime_by_month, Month, year)

############## crime by day
crime_by_day <- aggregate(year ~crime_type + day,
                           data = dat, FUN = length)
crime_by_day$day_num[crime_by_day$day == "Sunday"] <-1
crime_by_day$day_num[crime_by_day$day == "Monday"] <-2
crime_by_day$day_num[crime_by_day$day == "Tuesday"] <-3
crime_by_day$day_num[crime_by_day$day == "Wednesday"] <-4
crime_by_day$day_num[crime_by_day$day == "Thursday"] <-5
crime_by_day$day_num[crime_by_day$day == "Friday"] <-6
crime_by_day$day_num[crime_by_day$day == "Saturday"] <-7

crime_by_day <-crime_by_day[order(crime_by_day$day_num, decreasing = FALSE),]
crime_by_day$day <- factor(crime_by_day$day, levels=c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))

#############3 crime by hour
crime_by_hour <- aggregate(year ~crime_type + Hour,
                           data = dat, FUN = length)


##### plot crime related to time
plot.A <-ggplot(crime_by_year, aes(x = factor(year), y = Month, colour = crime_type)) + 
    geom_boxplot() + 
    geom_point(data = crime_by_year, aes(y = Month)) +
    geom_line(data = crime_by_year, aes(y = Month, group = crime_type)) + 
    theme_bw()
plot.A <-plot.A + theme(legend.position="none",
           axis.text.x  = element_text(angle=90, vjust=0.5, size=8))+
  xlab("") + ylab("Frequency")+labs(title="Year")
plot.B <-ggplot(crime_by_month, aes(x = factor(Month), y = year, colour = crime_type)) + 
    geom_boxplot() + 
    geom_point(data = crime_by_month, aes(y = year)) +
    geom_line(data = crime_by_month, aes(y = year, group = crime_type)) + 
    theme_bw()
plot.B <-plot.B + theme(legend.position="none",
           axis.text.x  = element_text(angle=90, vjust=0.5, size=8)) +
  xlab("") + ylab("")+labs(title="Month")

plot.C <-ggplot(crime_by_day, aes(x = factor(day), y = year, colour = crime_type)) + 
    geom_boxplot() + 
    geom_point(data = crime_by_day, aes(y = year)) +
    geom_line(data = crime_by_day, aes(y = year, group = crime_type)) + 
    theme_bw()
plot.C <-plot.C + theme(legend.position="none",
           axis.text.x  = element_text(angle=90, vjust=0.5, size=8)) +
  xlab("") + ylab("")+labs(title="Day")

plot.D <- ggplot(crime_by_hour, aes(x = factor(Hour), y = year, colour = crime_type)) + 
    geom_boxplot() + 
    geom_point(data = crime_by_hour, aes(y = year)) +
    geom_line(data = crime_by_hour, aes(y = year, group = crime_type)) + theme_bw()
plot.D <-plot.D + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8)) +   xlab("") + ylab("Frequency") +labs(title="Hour")


## plots are not even, so give 24hr more space than year, month and day
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 3)))
vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
print(plot.A, vp = vplayout(1, 1))  # key is to define vplayout
print(plot.B, vp = vplayout(1, 2))
print(plot.C, vp = vplayout(1, 3))
print(plot.D, vp = vplayout(2, 1:3))

#############################################################
```

**Geographical Look at Crime**  
Incorporating a third variable, location, crimes can be mapped with three criteria, place, time and type of crime. From the maps it is noticeable that crime clusters to different regions of the city.  These are geographical hotspots for crime. Clustering also is influenced by the time of year and type of crime.  

The first set of maps is the city of Philadelphia for violent and quality of life crimes commited in August of 2016.  There are overlaps to peaks of crime between the two types of crimes.  Quality is more concentrated through out the city, while violent has some concentration, but is otherwise spread out in sections.  

Violent crimes has had a year over year decline, while quality of life initially went down and than increased, but not to 2006.  To see if there was a location change in crime, plots were created to see how they changed over the years.  

In Sept 2010, SugarHouse Casino was open south of Fishtown.  There was concerned crime would increase.  From these graphs, the area right outside of the Casino does not show increase in crime, but the region Northern Kensington showed increase in crime activity since opending in Sept 2010.   


```{r, fig.width=3.5, fig.height=3.5, fig.path='Figs/'}

### subset data, this is important because the maps will take longer to graph if there is extra/too much data
library(ggmap)
library(ggplot2)

dat1 <-dat[ which(dat$Month =="08" &dat$year == 2016 & dat$crime_type =="Violent_crime"), ]
dat2 <-dat[ which(dat$Month =="08" &dat$year == 2016 & dat$crime_type =="Quality_of_Life"), ]
dat3 <-dat[ which(dat$Month =="02" &dat$year == 2016 & dat$crime_type =="Violent_crime"), ]
dat4 <-dat[ which(dat$Month =="02" &dat$year == 2016 & dat$crime_type =="Quality_of_Life"), ]
dat5 <-dat[ which(dat$Month =="08" &dat$year == 2006 & dat$crime_type =="Violent_crime"), ]
dat6 <-dat[ which(dat$Month =="08" &dat$year == 2006 & dat$crime_type =="Quality_of_Life"), ]
dat7 <-dat[ which(dat$Month =="02" &dat$year == 2006 & dat$crime_type =="Violent_crime"), ]
dat8 <-dat[ which(dat$Month =="02" &dat$year == 2006 & dat$crime_type =="Quality_of_Life"), ]
dat9 <-dat[ which(dat$Month =="08" & dat$crime_type =="Quality_of_Life"), ]
dat10 <-dat[ which(dat$Month =="08" & dat$crime_type =="Violent_crime"), ]


```

```{r, fig.width=3, fig.height=3, fig.path='Figs/'}

#######Graph with map set 1
m <- get_map("Philadelphia, PA",zoom=12,source = "google")
# another good source is "stamen"
g <- ggmap(m)

limits <-c(0,300)
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat1,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "blue",
                        limits=c(limits)) +labs(title="Aug Violent 2016")

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat2,geom="polygon", alpha=0.2, size = 2)
g + scale_fill_gradient(low = "white", high = "blue",
                        limits=c(limits)) +labs(title="Quality Aug 2016")


```


```{r, fig.width=1.8, fig.height=1.8, fig.path='Figs/', eval=FALSE, include=FALSE}

#m <- get_map("University City, Philadelphia, PA",zoom=14,source = "google")
m <- get_map("Philadelphia, PA",zoom=14,source = "google")
# another good source is "stamen"
g <- ggmap(m)



g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat1,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red",
                        limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
     legend.position="none")+labs(title="Violent Aug 2016")

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat2,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
     legend.position="none")+labs(title="Quality Aug 2016")

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat3,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Violent Feb 2016")


g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat4,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Quality Feb 2016")
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat5,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Violent August 2006")

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat6,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Quality Aug 2006")
#m <- get_map("Philadelphia, PA",zoom=12,source = "google")
# another good source is "stamen"
#g <- ggmap(m)
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat7,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Violent Feb 2006")
#m <- get_map("Philadelphia, PA",zoom=12,source = "google")
# another good source is "stamen"
#g <- ggmap(m)
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat8,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Quality Feb 2006")

limits <- c(50, 1500)

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat1,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red",
                        limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
     legend.position="none")+labs(title="Violent Aug 2016")

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat2,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
     legend.position="none")+labs(title="Quality Aug 2016")

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat3,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Violent Feb 2016")


g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat4,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Quality Feb 2016")
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat5,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Violent August 2006")

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat6,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Quality Aug 2006")
#m <- get_map("Philadelphia, PA",zoom=12,source = "google")
# another good source is "stamen"
#g <- ggmap(m)
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat7,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Violent Feb 2006")
#m <- get_map("Philadelphia, PA",zoom=12,source = "google")
# another good source is "stamen"
#g <- ggmap(m)
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat8,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "red", limits=c(limits))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none")+labs(title="Quality Feb 2006")
```


```{r, include=FALSE, eval=FALSE}
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat1,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red") +labs(title="Aug Violent 2016")

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat2,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red") +labs(title="Aug Violent 2016")
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat3,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red") +labs(title="Aug Violent 2016")
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat4,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red") +labs(title="Aug Violent 2016")
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat5,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red") +labs(title="Aug Violent 2016")
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat6,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red") +labs(title="Aug Violent 2016")
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat7,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red") +labs(title="Aug Violent 2016")
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat8,geom="polygon", alpha=0.2)
g + scale_fill_gradient(low = "white", high = "red") +labs(title="Aug Violent 2016")


```

```{r}

### graph on map with wrap of time: year, month, day, not hour too many
m <- get_map("Philadelphia, PA",zoom=12,source = "google")
# another good source is "stamen"
g <- ggmap(m)

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "blue")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
        labs(title="All Crime For Every Year")+
        facet_wrap(~ year)
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "blue")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
        labs(title="All Crime For Every Month")+
        facet_wrap(~ Month)
g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "blue")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
        labs(title="All Crime For Every Day")+
        facet_wrap(~ day)


g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat9,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "blue")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
        labs(title="Quality Crimes in Aug 2006:2016")+
        facet_wrap(~ year)
#g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat10,geom="polygon", alpha=0.2)
#g + scale_fill_gradient( low = "white", high = "blue")+
#  theme(axis.title.x=element_blank(),
#        axis.text.x=element_blank(),
#        axis.ticks.x=element_blank(),
#        axis.title.y=element_blank(),
#        axis.text.y=element_blank(),
#        axis.ticks.y=element_blank()) +
#        labs(title="Violence Crimers in Aug 2006:2016")+
#        facet_wrap(~ year)

m <- get_map("Fishtown, Philadelphia, PA",zoom=14,source = "google")
# another good source is "stamen"
g <- ggmap(m)

g <- g+stat_density2d(aes(x = Lon, y = Lat, fill=..level..), data=dat,geom="polygon", alpha=0.2)
g + scale_fill_gradient( low = "white", high = "blue")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
        labs(title="All Crimes 2006:2016")+
        facet_wrap(~ year)


```

```{r,fig.width=3, fig.height=3.5, include= FALSE, eval=FALSE}
par(mfrow = c(1,1))
upenn <- get_map("University City, Pennsylvania", zoom = 14)
PhiladelphiaMap <- ggmap(upenn, extent = "device") #, legend = "bottomleft")
PhiladelphiaMap +
  stat_density2d(
    aes(x = Lon, y = Lat, fill = ..level.., alpha = ..level..),
    size = 2, bins = 4, data = dat2,
    geom = "polygon"
  )

Chestnut <- get_map("Chestnut Hill, Philadelphia, Pennsylvania", zoom = 14)
PhiladelphiaMap2 <- ggmap(Chestnut, extent = "device") #, legend = "bottomleft")
PhiladelphiaMap2 +
  stat_density2d(
    aes(x = Lon, y = Lat, fill = ..level.., alpha = ..level..),
    size = 2, bins = 4, data = dat2,
    geom = "polygon"
  ) 


```

**Notes about Arrest Data**  
The data for the study was downloaded from Kaggle.  A website active in  predictive modeling competitions.  

Entries are not convictions, but reports generated from police officers, radio call personal and office support officers.  If crimes are not reported, they will not be in this anaysis.  If crimes are falsely reported they will also be included.  Homices produce  indisputable evidence, fluctations will not be based on priorities from Captains.  Quality of life crimes, on the other hand, can go unnotices and unreported.  This can create fluctuations in the data.        

Another area where the data is not precise is in the areas without addresses such as the Devil's pool.  This area in Fairmont park does not have a street address and it is a know place for incidents, medical and criminal happen.  Devil's pool does not have an address, so the closest address is used (Philly Magazine, July 26, 2015).

**Conclusion**  
From 2006 to 2016 crime in philadelphia has decreased.  There are more quality of life reports, but violent and other offenses are down.  The least amount of crime occurs in feb with the most in Aug.  Depending on the day, depends on the type of crime.  The same goes for hour, with the least amount of crime occuring in the earling morning between 4-7am.



